Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-12 methods_paper/fabl-paper.tex
added traceplots from one simulation

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.

@@

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics. Note that $\mathcal{P}$ and $\mathcal{R}$ fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}\}$ for our efficient posterior inference.

@@

With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\). Therefore we know that records $\i in r_{j_p}$  such that
\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Posterior calculations still attribute the appropriate weight to all records through the summary statistics \(H_p\), and \(H_{j_p}\). Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.

@@

>>>>

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. Open source software to use \texttt{fabl} in \texttt{R} is available through Github.

@@

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
where $r_{j_p} = \{i \in \bm{X_1} | (i, j) \in h_p\}$. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I((i, j) \in h_p) = ||r_{j_p}||$$
and
$$H_p =  \sum_{i, j} I((i, j) \in h_p) = \sum_{j} H_{j_p}$$
For convenience, denote this set of sufficient statistics $\mathcal{H} = \left\{\{H_{j_p}\}, \{H_{j_p}\}\right\}$As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics. Note that $\mathcal{P}$, $\mathcal{R}$, and $\mathcal{H}$  fully characterize the comparison matrix $\Gamma$ with no loss of information, so we proceed with $\tilde{\Gamma} = \{\mathcal{P}, \mathcal{R}, \mathcal{H}\}$ for our efficient posterior inference.

@@

With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\). Therefore we know that records $\i in r_{j_p}$  such that
\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be. Posterior calculations still attribute the appropriate weight to all records through the summary statistics in $\mathcal{H}$, and thus we can proceed with posterior inference through the memory reduced $\tilde{\Gamma}^{\text{SEI}} = \{\mathcal{P}, \mathcal{R}^{\text{SEI}}, \mathcal{H}\}$. For simplicty, we suppress ``SEI'' in future notation, and use $\tilde{\Gamma}$ without loss of generality.

@@

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_overlap_trace} 
		\caption{Representative traceplot of overlap between files from simulation study in \ref{accuracy}}\label{fig:sim_overlap_trace}
	\end{center}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_m_trace} 
		\caption{Representative traceplot of m parameter from simulation study in \ref{accuracy}}\label{fig:sim_m_trace}
	\end{center}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.6\textwidth]{../notes/figures/sim_u_trace} 
		\caption{Representative traceplot of u parameters from simulation study in \ref{accuracy}}\label{fig:sim_u_trace}
	\end{center}
\end{figure}
